---
title: "Geography 176A"
author: "[Kiki Takeuchi](https://github.com/ktakeuchi2)" 
subtitle: "Lab 04: Tesselations, Spatial Joins, and Point-in-Polygon"
output:
  html_document:
    theme: journal
---
*****

### Question 1 
##### Producing 5 Different Tesselations from CONUS

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(sf)
library(tidyverse)
library(USAboundaries)
library(USAboundariesData)
library(units)
library(readxl)
library(rmapshaper)
library(gghighlight)
library(knitr)
library(leaflet)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
counties = USAboundaries::us_counties() %>% 
  filter(!state_name %in% c("Hawaii", "Alaska", "Puerto Rico", "Guam")) %>% 
  st_transform(5070) %>% 
  st_as_sf()

county_cent = st_centroid(counties) %>% 
  st_union()
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
v_grid = st_voronoi(county_cent) %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  mutate(id = 1:n())

t_grid = st_triangulate(county_cent) %>% 
  st_cast() %>% 
  st_as_sf() %>% 
  mutate(id = 1:n())

sq_grid = st_make_grid(counties, n = 70) %>% 
  st_as_sf() %>% 
  mutate(id = 1:n())

hex_grid = st_make_grid(counties, n = 70, square = FALSE) %>% 
  st_as_sf() %>% 
  mutate(id = 1:n())
```
##### Simplifying the County Object

Original County Union:
```{r, message = FALSE, warning = FALSE, echo = FALSE}
counties_u = counties %>% 
  st_union()
plot(counties_u)
mapview::npts(counties_u)
```
Simplified Object:
```{r, message = FALSE, warning = FALSE, echo = FALSE}
counties_simp = rmapshaper::ms_simplify(counties_u, keep = 0.1)
plot(counties_simp)
mapview::npts(counties_simp)

v_grid = st_intersection(v_grid, counties_simp)
t_grid = st_intersection(t_grid, counties_simp)
```
Original: 3229 points, simplified object: 322 points, our simplified object removed 2907 points.
Consequences of doing this computationally are that we lose accuracy in the shape of our object, however, we do not need to spend as much time or data space on unnecessary points.

*****
##### Tesselations:

```{r, message = FALSE, warning = FALSE, echo = FALSE}
plot_tess = function(data, title) {
  ggplot() +
    geom_sf(data = data, fill = "white", col = "navy", size = 0.2) +
    theme_void() +
    labs(title = title, caption = paste("This Tesselation has:", nrow(data), "tiles")) +
    theme(plot.title = element_text(hjust = 0.5, color = "navy", face = "bold"))
}
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
plot_tess(counties, "Original County Data") +
  geom_sf(data = county_cent, col = "darkred", size = 0.2)

plot_tess(v_grid, "Voronoi Coverage") +
  geom_sf(data = county_cent, col = "darkred", size = 0.2)

plot_tess(t_grid, "Delaunay Triangulated Coverage") +
  geom_sf(data = county_cent, col = "darkred", size = 0.2)

plot_tess(sq_grid, "Square Coverage")

plot_tess(hex_grid, "Hexagonal Coverage")
```

*****

### Question 2
##### Writing a function to Summarize Tesselated Surfaces

```{r, message = FALSE, warning = FALSE, echo = FALSE}
sum_tess = function(data, text){
  area = st_area(data) %>% 
    units:: set_units("km^2") %>% 
    units::drop_units()
  tess_df = data.frame(text, nrow(data), mean(area), sd(area), sum(area))
  return(tess_df)}
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
tess_sum = bind_rows(
  sum_tess(counties, 'US Counties'),
  sum_tess(v_grid, 'Voronoi'),
  sum_tess(t_grid, 'Triangulated'),
  sum_tess(sq_grid, 'Square'),
  sum_tess(hex_grid, 'Hexagonal'))

knitr::kable(tess_sum,
             caption = "Tesselation Summary",
             col.names = c("Type", "Number of Features", "Mean Area (km^2)", "Standard Deviation", "Total Area (km^2)")) %>% 
   kableExtra::kable_styling("striped", full_width = TRUE)
```
Analyzing the tesselation summaries for each type of tesselation, it is notable that the Hexagonal Tesselation has the fewest number of features, and the largest total area. Another thing to note is that both square and hexagonal tesselations have 0 as a value for the standard deviation.

*****

### Question 3
##### Analyzing Data from US Army Corp of Engineers National Dam Inventory (NID)

```{r, message = FALSE, warning = FALSE, echo = FALSE}
dams <- read_excel("../data/NID2019_U.xlsx") %>%
  filter(!is.na(LONGITUDE), !is.na(LATITUDE)) %>%
  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326) %>%
  st_transform(5070)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
pip = function(points, polygon, group){
  st_join(polygon, points) %>%
    st_drop_geometry() %>%
    count(get(group)) %>%
    setNames(c(group, "n")) %>%
    left_join(polygon, by = group) %>%
    st_as_sf()
}
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
plot_pip = function(data, title){
  ggplot() +
    geom_sf(data = data, aes(fill = log(n)), col = NA, alpha = 0.9, size = 0.2) +
    scale_fill_viridis_c() +
    theme_void() +
    labs(title = title,
         caption = paste0(sum(data$n), "Dams")) +
    theme(legend.position = "none",
          plot.title = element_text(hjust = 0.5, color = "navy", face = "bold"))
}
```

```{r, message = FALSE, warning = FALSE}
counties_pip = pip(dams, counties, "geoid") %>% 
  plot_pip("Dams Per County")
v_pip = pip(dams, v_grid, "id") %>% 
  plot_pip("Dams Per Viroinoi Tesselation")
t_pip = pip(dams, t_grid, "id") %>% 
  plot_pip("Dams Per Triangulated Tesselation")
sq_pip = pip(dams, sq_grid, "id") %>% 
  plot_pip("Dams Per Square Coverage")
hex_pip = pip(dams, hex_grid, "id") %>% 
  plot_pip("Dams Per Hexagonal Coverage")
```

### Question 4
##### Analyzing Distribution and Function of US Dams

```{r, message = FALSE, warning = FALSE}
unique(dams$PURPOSES) %>%
  length

p_dams = dams %>%
  filter(grepl("P", dams$PURPOSES) == TRUE) 
p_pip = pip(p_dams, hex_grid, "id") %>% 
  plot_pip("Fire Protection Dams Over Hexagonal Tesselation") +
  gghighlight(n > mean(n) + sd(n))

i_dams = dams %>%
  filter(grepl("I", dams$PURPOSES) == TRUE)  
i_pip = pip(i_dams, hex_grid, "id") %>% 
  plot_pip("Irrigation Dams Over Hexagonal Tesselation") +
  gghighlight(n > mean(n) + sd(n))

f_dams = dams %>%
  filter(grepl("F", dams$PURPOSES) == TRUE)  
f_pip =  pip(f_dams, hex_grid, "id") %>% 
  plot_pip("Fish and Wildlife Dams Over Hexagonal Tesselation") +
  gghighlight(n > mean(n) + sd(n))

h_dams = dams %>%
  filter(grepl("H", dams$PURPOSES) == TRUE)
h_pip = pip(h_dams, hex_grid, "id") %>% 
  plot_pip("Hydroelectric Dams Over Hexagonal Tesselation") +
  gghighlight(n > mean(n) + sd(n))
```

*****


