---
title: "Geography 176A"
author: "[Kiki Takeuchi](https://github.com/ktakeuchi2)" 
subtitle: "Lab 03: Distances and the Border Zone"
output:
  html_document:
    theme: journal
---
*****

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(tidyverse)
library(sf)
library(units)
library(USAboundaries)
library(rnaturalearth)
library(gghighlight)
library(ggrepel)
library(knitr)
```


### Question 1
##### Creating Datasets for Boundaries of US States, USA, Canada, Mexico, and US Cities

In Part 1, we define the projection to the North America Equidistant Conic Projection. We also get boundaries for US states, the entire USA, Mexico, Canada, as well as US city locations.

```{r, message = FALSE, warning = FALSE}
eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
remotes::install_github("ropensci/USAboundaries")
remotes::install_github("ropensci/USAboundariesData")

conus = USAboundaries::us_states(resolution = "low") %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii")) %>% 
  st_transform(eqdc)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
remotes::install_github("ropenscilabs/rnaturalearthdata")

countries = rnaturalearth::countries110 
countries = countries %>%  
  st_as_sf() %>% 
  filter(admin %in% c("Mexico", "United States of America", "Canada")) %>% 
  st_transform(eqdc)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
cities = readr::read_csv("../data/uscities.csv") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  filter(!state_name %in% c("Hawaii", "Puerto Rico", "Alaska")) %>%
  st_transform(eqdc)
```

*****

### Question 2:
##### Calculating Distances of US cities to (1) the national US border, (2) the nearest state border, (3) the Mexican border, and (4) the Canadian border.

##### Distance to US Border (km)
```{r, message = FALSE, warning = FALSE, echo=FALSE}
conus_u = st_union(conus) %>%
  st_cast("MULTILINESTRING")

cities = cities %>% 
  mutate(dist_border = st_distance(cities, conus_u),
         dist_border = units::set_units(dist_border, "km"),
         dist_border = units::drop_units(dist_border))
  
farthest_us_border = cities %>%   
  select(city, state_name, dist_border) %>% 
  arrange(-dist_border) %>% 
  slice_max(dist_border, n = 5) %>% 
  st_drop_geometry()
  
knitr::kable(farthest_us_border,
             caption = "Cities Farthest from USA border",
             col.names = c("City", "State", "Distance to USA Border (km)")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE)
```

##### Distance to States (km)
```{r, message = FALSE, warning = FALSE, echo=FALSE}
conus_c = st_combine(conus) %>%
  st_cast("MULTILINESTRING")

cities = cities %>% 
  mutate(dist_state = st_distance(cities, conus_c),
         dist_state = units::set_units(dist_state, "km"),
         dist_state = units::drop_units(dist_state))

farthest_state <- cities %>%   
  select(city, state_name, dist_state) %>% 
  arrange(-dist_state) %>% 
  slice_max(dist_state, n= 5) %>%
  st_drop_geometry()


knitr::kable(farthest_state,
             caption = "Cities Farthest from State Borders",
             col.names = c("City", "State", "Distance to State Border (km)")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE)
```

##### Distance to Mexico (km)

```{r, message = FALSE, warning = FALSE, echo=FALSE}
mexico = countries %>% 
  filter(admin == "Mexico") %>% 
  st_cast("MULTILINESTRING")

cities = cities %>% 
  mutate(dist_mexico = st_distance(cities, mexico),
         dist_mexico = units::set_units(dist_mexico, "km"),
         dist_mexico = units::drop_units(dist_mexico))

farthest_mexico = cities %>%
  select(city, state_name, dist_mexico) %>% 
  arrange(-dist_mexico) %>% 
  slice_max(dist_mexico, n= 5) %>% 
  st_drop_geometry()
  
knitr::kable(farthest_mexico,
             caption = "Cities Farthest from Mexican Border",
             col.names = c("City", "State", "Distance to Mexican Border (km)")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE)
```

##### Distance to Canada (km)
```{r, message = FALSE, warning = FALSE, echo = FALSE}
canada = countries %>% 
  filter(admin == "Canada") %>% 
  st_cast("MULTILINESTRING")

cities = cities %>% 
  mutate(dist_canada = st_distance(cities, canada),
         dist_canada = units::set_units(dist_canada, "km"),
         dist_canada = units::drop_units(dist_canada))

farthest_canada = cities %>%
  select(city, state_name, dist_canada) %>% 
  arrange(-dist_canada) %>% 
  slice_max(dist_canada, n= 5) %>%
  st_drop_geometry()

knitr::kable(farthest_canada,
             caption = "Cities Farthest from Canadian Border",
             col.names = c("City", "State", "Distance to Canadian Border (km)")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE)
```

*****

### Question 3
##### Visualization

```{r, message = FALSE, warning = FALSE, echo=FALSE}
cities_top10 = cities %>% 
  slice_max(population, n = 10)

ggplot() +
  geom_sf(data = countries, fill = "gray80") +
  geom_sf(data = conus_u, color = "blue", cex = 1) +
  geom_sf(data = conus_c, color = "black") +
  geom_sf(data = cities_top10, aes(size = population), color = "red") +
  theme_linedraw() +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "slategray1"),
        legend.position = "none") +
  labs(title = "Top 10 Largest US Cities")
```


```{r, message = FALSE, warning = FALSE, echo = FALSE}
far_us_border = cities %>% 
  select(city, state_name, dist_border) %>% 
  arrange(-dist_border) %>% 
  slice_max(dist_border, n = 5)

ggplot() +
  geom_sf(data = conus, fill = NA) +
  geom_sf(data = cities, aes(col = dist_border), size = 0.1) +
  scale_color_gradient(low = "grey", high = "darkblue") +
  geom_sf(data = far_us_border, col = "red", size = 1) +
  ggrepel::geom_label_repel(data = far_us_border,
                            aes(label = city, geometry = geometry),
                            stat = "sf_coordinates",
                            size = 4) +
  labs(title = "Distance of US Cities to US Border",
       col = "Distance to US Border (km)") +
  ggthemes::theme_map() 
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
far_state = cities %>% 
  select(city, state_name, dist_state) %>% 
  arrange(-dist_state) %>% 
  slice_max(dist_state, n= 5)

ggplot() +
  geom_sf(data = conus, fill = NA) +
  geom_sf(data = cities, aes(col = dist_state), size = 0.2) +
  geom_sf(data = far_state, col = "navyblue", size = 1) +
  scale_color_gradient(low = "grey", high = "darkred") +
  ggrepel::geom_label_repel(data = far_state,
                            aes(label = city, geometry = geometry),
                            stat = "sf_coordinates",
                            size = 4) +
  labs(title = "Distance of US Cities to State Borders",
       col = "Distance to US Border (km)") +
  ggthemes::theme_map()
```

##### 3.4
```{r, message = FALSE, warning = FALSE, echo = FALSE}
eqdistance = cities %>% 
  mutate(ab_distance = abs(dist_mexico - dist_canada)) %>% 
  filter(ab_distance <= 100)

eqdistance_5 = eqdistance %>% 
  slice_max(population, n = 5)

ggplot() +
  geom_sf(data = conus, fill = NA) +
  geom_sf(data = eqdistance, aes(col = ab_distance), size = 0.2) +
  geom_sf(data = eqdistance_5, col = "red", size = 2) +
  labs(title = "US Cities Equidistant From Mexican and Canadian Borders") +
  ggthemes::theme_map() +
  ggrepel::geom_label_repel(data = eqdistance_5,
                            aes(label = city, geometry = geometry),
                            stat = "sf_coordinates",
                            size = 3) +
  theme(legend.position = "none")
```

*****

### Question 4
##### Real-World Application: Quantifying and Mapping the "Border Zone"
Information can be found from [ACLU](https://www.aclu.org/other/constitution-100-mile-border-zone).

```{r, message = FALSE, warning = FALSE, echo = FALSE}
zone = cities %>% 
  mutate(totalpop = sum(population)) %>% 
  filter(dist_border <= 160) %>% 
  summarise(zone_cities = n(), zone_pop = sum(population), percent = (zone_pop / totalpop) * 100) %>% 
  slice(n = 1) %>% 
  st_drop_geometry()

knitr::kable(zone, 
             align = 'l',
             caption = "100 Mile Border Zone",
             col.names = c("Number of Cities", "Population Within Zone", "Percent of Total Population")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE)
```
The findings in this table support the claim made by ACLU, that 2/3 of the US population live in the "Border Zone".

```{r, message = FALSE, warning = FALSE, echo = FALSE}
zone_10 = cities %>% 
  filter(dist_border <= 160) %>% 
  group_by(state_name) %>% 
  slice_max(population, n = 1)

ggplot() +
  geom_sf(data = conus, fill = NA) +
  geom_sf(data = cities, col = "black", size = 0.1) +
  geom_sf(data = zone_10, col = "navyblue", size = 1) +
  scale_color_gradient(low = "orange", high = "darkred") +
  gghighlight(dist_border <= 160) +
  labs(title = "Border Zone",
       subtitle = "100 Mile Zone / Danger Zone") +
  ggthemes::theme_map() +
  ggrepel::geom_label_repel(data = zone_10,
                            aes(label = city, geometry = geometry),
                            stat = "sf_coordinates",
                            size = 3) +
  theme(legend.position = "none") +
  ggsave(file = "../img/lab03_us_border_zone.png")
```

*****

Constitutional rights protected by the Fourth Constitutional Amendment no longer apply fully at borders. The border zone of 100 miles (~160 km) from external boundaries contains 28,145 cities. 259 million people live in these border zones. Compared to the 397 million people living in the US, 65 % of the US population, or about two thirds, is living in these border zones. This is accurate to the ACLU article estimate.

*****
