---
title: "Geography 176A"
author: "Kiki Takeuchi"
subtitle: "Lab 05: Rasters and Remote Sensing"
output:
  html_document:
  theme: journal
---

### Question 1
Identifying the Area of Interest

```{r, message = FALSE, warning = FALSE}
library(raster)
library(tidyverse)
library(getlandsat)
library(sf)
library(mapview)
library(osmdata)

bb = read_csv("../data/uscities.csv") %>%
  filter(city == "Palo") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(5070) %>%
  st_buffer(5000) %>%
  st_bbox() %>%
  st_as_sfc() %>%
  st_as_sf()

bbwgs = bb %>%
  st_transform(4326)

```

### Question 2
Identifying, Downloading, and Caching Images

```{r, message = FALSE, warning = FALSE}
meta = read_csv("../data/palo-flood.csv")

files = lsat_scene_files(meta$download_url) %>% 
  filter(grepl(paste0("B", 1:6, ".TIF$", collapse = "|"), file)) %>% 
  arrange(file) %>% 
  pull(file)

st = sapply(files, lsat_image)

b = stack(st) %>% 
  setNames(c(paste0("band", 1:6)))

cropper = bbwgs %>% 
  st_transform(crs(b))

r = crop(b, cropper)
```

##### Dimensions, Coordinate Reference System, and Resolution of stacked images

```{r, message = FALSE, warning = FALSE}
dim(b)
crs(b)
res(b)

dim(r)
crs(r)
res(r)
```


### Question 3
Making RGB plots of these images.

```{r, message = FALSE, warning = FALSE}
par(mfrow = c(2, 2))
plotRGB(r, r = 4, g = 3, b = 2)
plotRGB(r, r = 5, g = 4, b = 3)
plotRGB(r, r = 6, g = 5, b = 4)
plotRGB(r, r = 5, g = 7, b = 1)


par(mfrow = c(2, 2))
plotRGB(r, r = 4, g = 3, b = 2, stretch = "lin")
plotRGB(r, r = 5, g = 4, b = 3, stretch = "lin")
plotRGB(r, r = 6, g = 5, b = 4, stretch = "lin")
plotRGB(r, r = 5, g = 7, b = 1, stretch = "lin")


par(mfrow = c(2, 2))
plotRGB(r, r = 4, g = 3, b = 2, stretch = "hist")
plotRGB(r, r = 5, g = 4, b = 3, stretch = "hist")
plotRGB(r, r = 6, g = 5, b = 4, stretch = "hist")
plotRGB(r, r = 5, g = 7, b = 1, stretch = "hist")
```

Applying a color stretch helps make differences in landscapes more visibly clear to assist in identifying patterns or features.

### Question 4
Assessments of Surface Water Features Using A Threshold Value.

Raster Algebra for 5 Formulas
```{r, message = FALSE, warning = FALSE}
ndvi = (r$band5 - r$band4) / (r$band5 + r$band4)
ndwi = (r$band3 - r$band5) / (r$band3 + r$band5)
mndwi = (r$band3 - r$band6) / (r$band3 + r$band6)
wri = (r$band3 + r$band4) / (r$band5 + r$band6)
swi = 1 / sqrt(r$band2 - r$band6)

sw_rasters = stack(ndvi, ndwi, mndwi, wri, swi) %>% 
  setNames(c("ndvi", "ndwi", "mndwi", "wri", "swi"))

palette = colorRampPalette(c("blue", "white", "red"))
plot(sw_rasters, col = palette(256))
```

The images are similar in that they highlight the river region as well as other surface water areas. They also all use the the same blue, white, and red color scheme that we created in our palette. They differ in that the surface water areas are not all highlighted in the same color, and some images direct more attention towards the landscape rather than water regions. The final plot, the simple water index, highlights only the surface water captured in the image.


Raster Thresholding
```{r, message = FALSE, warning = FALSE}
thresh_ndvi = function(x){ifelse(x <= 0, 1, NA)}
thresh_ndwi = function(x){ifelse(x >= 0, 1, NA)}
thresh_wri = function(x){ifelse(x >= 1, 1, NA)}
thresh_swi = function(x){ifelse(x <= 5, 1, NA)}

flood_ndvi = calc(ndvi, thresh_ndvi)
flood_ndwi = calc(ndwi, thresh_ndwi)
flood_mndwi = calc(mndwi, thresh_ndwi)
flood_wri = calc(wri, thresh_wri)
flood_swi = calc(swi, thresh_swi)

flood = stack(flood_ndvi, flood_ndwi, flood_mndwi, flood_wri, flood_swi) %>% 
  setNames(c("ndvi", "ndwi", "mndwi", "wri", "swi"))

plot(flood, col = "blue")
flood <- na.omit(flood)
```

### Question 5
Supervised and Unsupervised Classification

```{r, message = FALSE, warning = FALSE}
set.seed(09072020)
dim(r)

values = values(r)
idx = which(!is.na(values))
v = na.omit(values)
vs = scale(v)

library(stats)

k12 = kmeans(vs, centers = 12)
kmeans_raster = flood$ndvi
values(kmeans_raster) = k12$cluster
plot(kmeans_raster)

k11 = kmeans(vs, centers = 11)
kmeans_raster = flood$ndvi
values(kmeans_raster) = k11$cluster
plot(kmeans_raster)

k10 = kmeans(vs, centers = 10)
kmeans_raster = flood$ndvi
values(kmeans_raster) = k10$cluster
plot(kmeans_raster)

k13 = kmeans(vs, centers = 13)
kmeans_raster = flood$ndvi
values(kmeans_raster) = k13$cluster
plot(kmeans_raster)

k9 = kmeans(vs, centers = 9)
kmeans_raster = flood$ndvi
values(kmeans_raster) = k9$cluster
plot(kmeans_raster)

flood_values = getValues(flood_ndvi)
kmeans_values = getValues(kmeans_raster)
table = table(flood_values, kmeans_values)
which.max(table)

mask = function(x){ifelse(x == 5, 1, NA)}
flood_mask = calc(kmeans_raster, mask)

flood = addLayer(flood, flood_mask) 
```

### Question 6

```{r, message = FALSE, warning = FALSE}
s1 = cellStats(flood$ndvi, stat = 'sum') * 30^2
s2 = cellStats(flood$ndwi, stat = 'sum') * 30^2
s3 = cellStats(flood$mndwi, stat = 'sum') * 30^2
s4 = cellStats(flood$swi, stat = 'sum') * 30^2
s5 = cellStats(flood$wri, stat = 'sum') * 30^2
s6 = cellStats(flood_mask, stat = 'sum') * 30^2
comparison = cbind(method = c("ndvi", "ndwi", "mndwi", "wri", "swi", "kmeans_mask"), cell_area = c(s1, s2, s3, s4, s5, s6))


knitr::kable(comparison, 
             caption = "Total Area of Flooded Cells",
             col.names = c("Band", "Total Flooded Cells")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE)

total_flood = calc(flood, sum)
```

plot(total_flood, col = RColorBrewer::brewer.pal(9, "Blues"))

plot shows up in R but shows up as empty plot in Rmd.

total_flood <- na.omit(total_flood)
flood = addLayer(flood, total_flood)
mapview(total_flood)

*****

