

```{r}
covid %>% 
  filter(state == state.of.interest) %>% 
  group_by(date) %>% 
  mutate(newCases = cases - lag(cases)) %>% 
  ungroup() %>%   
  kable(cases = sum(cases, na.rm = FALSE), caption = "Most Cumulative Cases") %>% 
  
  kable(newCases, caption = "Most New Cases")
  knitr::kable(cases = sum(cases, na.rm = FALSE), caption = "Most Cumulative Cases", head(5))
  knitr::kable(newCases, caption = "Most New Cases", head(5))
  
  
```

```{r}
dat0 = covid %>% 
  filter(state == "California") %>% 
  group_by(county) %>% 
  summarize(sum_cases = sum(cases)) %>% 
  ungroup() %>% 
  filter(date == max(date)) %>% 
  slice_max(sum_cases, n = 5) %>% 
  select(county, sum_cases) %>% 
  knitr::kable(sum_cases,
               caption = "Most Cumulative Cases",
               col.names = c("County", "Cases")) %>% 
    kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)
``` 


#### Question 2

install.packages("zoo")
library(zoo)

covid %>% 
group_by(state) %>% 
filter(state == "New York", "California", "Louisiana", "Florida") %>% 
group_by(date) %>% 
summarize(cases = sum(cases)) %>% 
mutate(newCases = cases - lag(cases),
  roll7 = rollmean(newCases, 7, fill = NA, align = "right")) %>% 
  ggplot(aes (x = date)) +
  geom_col(aes(y = newCases), col = state, fill = state)) +
  geom_line(aes(y = roll7, col = state, size = 1)) +
  facet_wrap(date~state, scales = "free_y") +
  labs(title = "Daily New Cases and 7-Day Rolling Mean for New York, California, Louisiana, and Florida",
  x = "Date",
  y = "Cases",
  subtitle = ")
  
  



homes = read_csv("data/landdata-states.csv")

library(readxl)
pop <- read_excel("data/PopulationEstimates.xls", skip = 2) %>% 
  select(pop19 = POP_ESTIMATE_2019, state = State, county = Area_Name) %>% 
  group_by(state) %>% 
  slice_max(pop19, n=1)
  

summary(pop)
dim(pop)
names(pop)
head(pop)


homes %>% 
  filter(State %in% c("CA", "TX")) %>% 
  ggplot(aes(x = Year, y = Home.Value)) +
  geom_line(aes(color = State)) +
  labs(title = "Title") +
  ggthemes::theme_excel()
  
homes$Date




```{r}
library(readxl)
PopulationEstimates <- read_excel("../data/PopulationEstimates.xls", 
    skip = 2) 
  
pop = PopulationEstimates %>% 
  select(fips = FIPStxt, pop19 = POP_ESTIMATE_2019, state = State, county = Area_Name) %>% 
  group_by(state) %>% 
  slice_max(pop19, n=1) %>% 
  right_join(dat, pop, by = "fips") %>%  
  group_by(county) %>% 
  mutate(new_cases_per_cap = sum(most_new_cases)/pop19) %>% 
  ungroup()
```

```{r, message=FALSE, warning=FALSE}
most_cases_per_cap = pop %>% 
  group_by(county) %>% 
  summarize(cases_per_cap = sum(cases)/pop19) %>% 
  slice_max(cases_per_cap, n = 5) %>% 
  select(county, cases_per_cap)

knitr::kable(most_cases_per_cap,
             caption = "Most Cumulative Cases Per Capita",
             col.names = c("County", "Cases Per Capita")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE)
```

```{r}
most_new_cases_per_cap = pop %>% 
  slice_max(new_cases_per_cap, n = 5) %>% 
  select(county, new_cases_per_cap)

knitr::kable(most_new_cases_per_cap,
             caption = "Most New Cases Per Capita",
             col.names = c("County", "New Cases Per Capita")) %>% 
  kableExtra::kable_styling("striped", full_width = TRUE)
```

```{r}
pop %>% 
  filter(date == max(date)-13) %>% 
  group_by(county, pop19) %>% 
  summarize(tot_cases = sum(cases) / 100000) %>% 
  mutate(newCases = cases - lag(cases)) %>% 
  summarize(new_Cases = sum(newCases) / 100000) %>% 
  ungroup() %>% 
  count(new_Cases)
```

The total number of cases is , the total number of new cases is , and the total number of safe counties, where there are less than 100 new cases per 100,000 residents over the past 14 days is .
*****

#### Question 2: COVID Cases in New York, California, Louisiana, and Florida

```{r}
library(zoo)

```


